### 2014-08-12 proposal/basic-prop.tex
Added basic research proposal

\documentclass[a4paper]{article}
\usepackage[cm]{fullpage}

\title{Honours 2014: Research Proposal}
\author{Tianyu Pu}

\date{Semester 1, 2014}

\begin{document}

\maketitle

\section{Title}
The research question is still being refined, but the title will be something
along the lines of ``Using data mining techniques to predict the length of stay
or mortality outcome of patients admitted to a major trauma centre".

\section{Supervisor}
Irena Koprinska, \texttt{irena.koprinska@sydney.edu.au} \\
School of Information Technologies, 4E

\section{Abstract}
The aim of the study is to derive a model using various artificial intelligence
and data mining techniques to predict the length of stay and mortality outcomes
of patients that are admitted to a major trauma centre. This model will be
internally validated using data from the Royal Prince Alfred Hospital. A mobile
application will be developed to assist planning and decision-making for
doctors in the trauma centre. There will also be some room to investigate
whether such a model can be generalised or used by other medical departments to
assist in their resource allocation and treatment preparation.

\section{Desk No.}
4E14

\bibliographystyle{plain}
\bibliography{references}
\nocite{*}

\end{document}