### 2014-08-12 poster/myposter.tex
Added poster template and skeleton files

\documentclass[portrait]{usydposter}
\usepackage{LI}
\usepackage{xspace}

\newcommand{\acronym}[1]{\textsc{#1}\xspace}
\newcommand{\cf}[1]{\mbox{$\it{#1}$}}
\newcommand{\todo}[1]{{\color{red} #1}}

\newcommand{\ngram}{n-gram\xspace}
\newcommand{\ngrams}{{\ngram}s\xspace}
\newcommand{\candc}{C\&C\xspace}
\newcommand{\ccgbank}{CCGBank\xspace}
\newcommand{\ccg}{\acronym{ccg}}
\newcommand{\cky}{\acronym{cky}}
\newcommand{\nlp}{\acronym{nlp}}
\newcommand{\np}{\acronym{np}}
\newcommand{\pos}{\acronym{pos}}
\newcommand{\wsj}{\acronym{wsj}}

\flushbottom

\title{CCG Parsing with One Structure per n-gram}
\author{Tim Dawborn}

\begin{document}
\makeheader

\begin{multicols}{3}

% =============================================================================
\section{Heading RGB 206, 17, 38 -- Arial 28}
\noindent This is some text here...

\begin{itemize}
  \item Parsing is the process of deriving the grammatical structure for a given sentence
  \item Natural language parsing is currently slow, with the best algorithms running in $O(n^3)$ with respect to the length of the sentence
  \item There is an inherent \emph{redundancy} in natural languages where certain common phrases (or \ngrams) appear frequently in text, each time with the same grammatical structure
  \item We explore the idea of exploiting this redundancy through the construction and \emph{memoisation} of the parse structures for these frequent \ngrams
\end{itemize}

% =============================================================================
\section{Background}
\begin{itemize}
  \item \cky is a $O(n^3)$ parsing algorithm which uses a chart data structure, illustrated in Figure~1. A chart is a hierarchical triangular structure  which stores the parse trees for all substrings of the sentence being parsed
\end{itemize}

\begin{figure}
  \pgfimage[width=0.7\columnwidth]{images/chart}
  \caption{An illustration of the chart data structure used in parsing algorithms such as \cky}
\end{figure}

\begin{itemize}
  \item Combinatory Categorial Grammar (\ccg) \cite{steedman:00} is a \emph{lexicalised grammar formalism}
  \item Each word in a sentence is assigned a composite object that reflects its role within the sentence
  \item The following is an example derivation using \ccg, where the lexical categories for each token are shown
\end{itemize}

\begin{center}
  \small
  \deriv{4}{
    \rm I & \rm like & \rm the & \rm cat \\
    \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S[dcl]\bs NP)/NP}} & \mc{1}{\cf{NP[nb]/N}} & \mc{1}{\cf{N}} \\
    & & \fapply{2} \\
    & & \mc{2}{\cf{NP[nb]}} \\
    & \bapply{3}\\
    & \mc{3}{\cf{S[dcl]\bs NP}}\\
    \bapply{4}\\
    \mc{4}{\cf{S[dcl]}}\\
  }
\end{center}

% =============================================================================
\section{Problems with Pizza}
\emph{Syntactic ambiguity} is a problem when parsing, and is a particular problem for our one structure per \ngram problem. Consider the sentence

\begin{center}
  \small
  \deriv{5}{
    \rm I & \rm ate & \rm pizza & \rm with & \rm anchovies \\
    \uline{1} & \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S\bs NP)/NP}} & \mc{1}{\cf{NP}} & \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP}} \\
    & & & \fapply{2} \\
    & & & \mc{2}{\cf{NP\bs NP}} \\
    & & \bapply{3}\\
    & & \mc{3}{\cf{NP}}\\
    & \fapply{4} \\
    & \mc{4}{\cf{S\bs NP}} \\
    \bapply{5}\\
    \mc{5}{\cf{S}}\\
  }
\end{center}

and compare it to the sentence

\begin{center}
  \small
  \deriv{5}{
    \rm I & \rm ate & \rm pizza & \rm with & \rm cutlery \\
    \uline{1} & \uline{1} & \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(S\bs NP)/NP}} & \mc{1}{\cf{NP}} & \mc{1}{\cf{((S\bs NP)\bs(S\bs NP))/NP}} & \mc{1}{\cf{NP}} \\
    & \fapply{2} & \fapply{2} \\
    & \mc{2}{\cf{S\bs NP}} & \mc{2}{\cf{(S\bs NP)\bs(S\bs NP)}} \\
    & \bapply{4} \\
    & \mc{4}{\cf{S\bs NP}} \\
    \bapply{5}\\
    \mc{5}{\cf{S}}\\
  }
\end{center}

These two sentences differ by only one word, yet they have two completely different derivations due to the different roles the word \textrm{with} plays in each sentence.

The word \textrm{with} thus is syntactically ambiguous. Without knowing the surrounding context, it is unknown which grammatical role the word is undertaking.

% =============================================================================
\section{Non-constituent-forming \ngrams}
Memoisation of \ngrams which do not form constituents can be achieved through the use of {\ccg}'s \emph{composition} and \emph{type raising} combinators.

\begin{center}
  \small
  \deriv{3}{
    \rm of & \rm the & \rm potato \\
    \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} & \mc{1}{\cf{N}} \\
    & \fapply{2}\\
    & \mc{2}{\cf{NP}} \\
    \fapply{3}\\
    \mc{3}{\cf{NP\bs NP}}\\
  }
\end{center}

\textrm{The} is joined to \textrm{potato} before \textrm{of} can be joined with \textrm{the}. Instead, we could construct the following derivation and insert it into the pre-constructed database:

\begin{center}
  \small
  \deriv{2}{
    \rm of & \rm the \\
    \uline{1} & \uline{1} \\
    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} \\
    \fcomp{2} \\
    \mc{2}{\cf{(NP\bs NP)/N}} \\
  }
\end{center}

Here we use \ccg forwards composition to combine \textrm{of} and \textrm{the} into a constituent-forming analysis. This chart structure could be reused to construct a span of the original phrase in the following manner:

\begin{center}
  \small
  \deriv{3}{
    \rm of & \rm the & \rm potato \\
    \uline{1} & \uline{1} & \uline{1} \\
    \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{1}{\cf{NP/N}} & \mc{1}{\cf{N}} \\
    \fcomp{2} & \\
    \mc{2}{\cf{(NP\bs NP)/N}} & \\
    \fapply{3}\\
    \mc{3}{\cf{NP\bs NP}}\\
  }
\end{center}

One situation where this technique does not work is is with \emph{prepositional phrase attachment}. The correct \ccg derivation for the phrase \textrm{on the king of England} is

\begin{center}
  \small
  \deriv{6}{
    \rm X & \rm on & \rm the & \rm king & \rm of & \rm England \\
    \uline{1} & \uline{1} & \uline{2} & \uline{2}  \\
    \mc{1}{\cf{NP}} & \mc{1}{\cf{(NP\bs NP)/NP}} & \mc{2}{\cf{NP[nb]}} & \mc{2}{\cf{NP\bs NP}} \\
    & & \bapply{4}\\
    & & \mc{4}{\cf{NP}}\\
    & \fapply{5} \\
    & \mc{5}{\cf{NP\bs NP}} \\
    \bapply{6} \\
    \mc{6}{\cf{NP}} \\
  }
\end{center}

If we were to use in this example the same forward composed derivation of the bigram \textrm{of the} as described earlier for the bigram \textrm{on the}, the wrong analysis would be constructed.

\begin{center}
  \small
  \deriv{6}{
    \rm X & \rm on & \rm the & \rm king & \rm of & \rm England \\
    \uline{1} & \uline{2} & \uline{1} & \uline{2}  \\
    \mc{1}{\cf{NP}} & \mc{2}{\cf{(NP\bs NP)/N}} & \mc{1}{\cf{N}} & \mc{2}{\cf{NP\bs NP}} \\
    & \fapply{3} & & \\
    & \mc{3}{\cf{NP\bs NP}} & & \\
    \bapply{4} & & \\
    \mc{4}{\cf{NP}} & & \\
    \bapply{6} \\
    \mc{6}{\cf{NP}} \\
  }
\end{center}

% =============================================================================
\section{Implementation}
The \candc parser \cite{clark-curran:07} is a state of the art \ccg parser which uses \cky. The \candc parser was used to memoise constituent-forming \ngrams into databases (Figure~2). When parsing sentences in the future, the memoised databases can be used to insert pre-constructed parses straight into the chart (Figure~3).

\begin{figure}
  \pgfimage[width=0.5\columnwidth]{images/prebuilt1}
  \caption{When creating the trigram database, a trigram is added to the database if it forms a constituent in the chart}
  \label{fig:prebuilt1}
\end{figure}

\begin{figure}
  \pgfimage[width=0.7\columnwidth]{images/prebuilt2}
  \caption{Illustration of using the \ngram databases. The trigram {\tt B~C~D} is loaded from the pre-constructed database, and blocks the corresponding cells}
\end{figure}

% =============================================================================
\section{Results}
Empirical results for the memoisation of constituent-forming \ngrams can be seen in the table below.

\begin{itemize}
  \item Different grammar models were used over varying sized \ngrams, with the  parsing time, F-score, and coverage figures being recorded
  \item No statistically significant change was observed in the F-scores or parsing times
  \item These results backup the theoretical hypothesis that persisting only constituent-forming \ngrams would not be enough to get the one structure per \ngram idea to work in practice
\end{itemize}

\begin{table}
  \small
  \begin{tabular}{cc||c|cc}
    \hline
    Model & & Baseline & 2-gram & 3-gram \\
    \hline \hline
           &  Time  & 82.82   & 82.83   & 82.57  \\
    \wsj   &   LF   & 85.26   & 85.26   & 85.26  \\
    derivs &   UF   & 92.01   & 92.01   & 92.01  \\
           &   Cov  & 98.64   & 98.64   & 98.64  \\
%%%    \hline
%%%           &  Time  & 84.15   & 83.83   & 83.87  \\
%%%    \wsj   &   LF   & 87.40   & 87.40   & 87.40  \\
%%%    hybrid &   UF   & 93.10   & 93.10   & 93.10  \\
%%%           &   Cov  & 98.64   & 98.64   & 98.64  \\
%%%    \hline
%%%           &  Time  & 84.10   & 83.96   & 84.26  \\
%%%    \np    &   LF   & 83.60   & 83.60   & 83.60  \\
%%%    derivs &   UF   & 90.48   & 90.48   & 90.48  \\
%%%           &   Cov  & 98.54   & 98.54   & 98.54  \\
%%%    \hline
%%%           &  Time  & 84.82   & 84.88   & 85.32  \\
%%%    \np    &   LF   & 85.88   & 85.88   & 85.88  \\
%%%    hybrid &   UF   & 91.63   & 91.63   & 91.63  \\
%%%           &   Cov  & 98.54   & 98.54   & 98.54  \\
    \hline
  \end{tabular}
  \caption{The speed versus performance trade-off for varying sized \ngrams on \ccgbank 02 to 21}
  \label{tab:results}
\end{table}

%%%Experiments were also conducted using only non-constituent-forming \ngrams over the same datasets. However the outcome of these experiments were the same; no statistically significant change was observed.

% =============================================================================
\section{Conclusion}
Thorough analysis of this one structure per \ngram idea using \ccg as a grammar formalism combined with a set of empirical results has shown that this one structure per \ngram idea does not work in the general when using token based \ngrams.

% =============================================================================
\section{Future work}
\begin{itemize}
  \item Recent work has been undertaken to explore the one structure per \ngram idea using \ngrams constructed of supertags instead of tokens
  \item The initial results for this new direction are positive, showing show a 10\% parser speedup being achieved with minimal loss of F-score
\end{itemize}

% =============================================================================
\references
\bibliography{references}

\end{multicols}
\end{document}