### 2014-08-30 presentation/example.tex
Added template for hons presentation

\documentclass{beamer}
\usepackage{graphicx}
\usepackage{lmodern}

% Load the usyd template
\input{usyd-beamer/usyd-beamer.tex}

\begin{document}

\usydtitleframe{Title}{Subtitle}{Author information}

\usydtitleframe
{Title goes here, this one is heaps long and wraps around}
{
	\vspace{3mm} % we want a little bit more room...
	Subtitle goes here, this is pretty long too hey? But not quite as long...
} {
	\textbf{Firstname Lastname} \hfill Student \\
	SID: 123 456 789 \\
	\vspace{3mm}
	\textbf{Firstname Lastname} \hfill Supervisor \\
	%\vspace{-12mm} % You might need to play here to get it positioned nicely...
}

\section{First Section}
\subsection{Subsection}
\begin{frame}{A sample slide}

A displayed formula:

\[ \int_{-\infty}^\infty e^{-x^2} \, dx = \sqrt{\pi} \]

An itemized list:

\begin{itemize}
  \item itemized item 1
  \item itemized item 2
  \item itemized item 3
\end{itemize}

\begin{theorem}
  In a right triangle, the square of hypotenuse equals
  the sum of squares of two other sides.
\end{theorem}

\end{frame}

\end{document}