### 2014-09-07 thesis/introduction/introduction.tex
Added structure of introduction

\chapter{Introduction}

\section{Background}

\subsection{Problem Statement}

\subsection{Previous Work}

\section{Contributions}