### 2014-09-22 thesis/thanks/thanks.tex
Added acknowledgements

\chapter*{Acknowledgements}

I would like to thank my supervisor, Dr Irena Koprinska, for her unwavering
support throughout the year. Irena, thank you for always being willing to
listen, and for raising points that inspired me to pursue new directions in
my research.

I would also like to thank Dr Michael Dinh from the Royal Prince Alfred
Hospital, without whom the research in this thesis would not have been
possible. Thank you,
Michael, for providing access to the data we needed for this study, and for
patiently answering all my questions about the data set and your previous
experiments.

Thank you to my close friends Daniel Buckmaster, Stephen Caraher and Christine
Au Yeung for the selfless encouragement, help and sanity checks that have
made this year-long journey possible.

And finally, but no less significant -- to my parents Caixiu and Ning, and
my sister Lily -- thank you for tirelessly supporting me throughout this year
and in everything that I do.