### 2014-09-22 thesis/conclusion/conclusion.tex
Updated heading of conclusion to 'conclusions'

\chapter{Conclusions} \label{chap:conclusion}

The conclusion goes here.