### 2014-09-26 poster/myposter.tex
Updated poster template to the latest style #14

\documentclass[landscape]{usydposter}
\usepackage{xspace}

\newcommand{\acronym}[1]{\textsc{#1}\xspace}
\newcommand{\cf}[1]{\mbox{$\it{#1}$}}
\newcommand{\todo}[1]{{\color{red} #1}}

\newcommand{\ngram}{n-gram\xspace}
\newcommand{\ngrams}{{\ngram}s\xspace}
\newcommand{\candc}{C\&C\xspace}
\newcommand{\ccgbank}{CCGBank\xspace}
\newcommand{\ccg}{\acronym{ccg}}
\newcommand{\cky}{\acronym{cky}}
\newcommand{\nlp}{\acronym{nlp}}
\newcommand{\np}{\acronym{np}}
\newcommand{\pos}{\acronym{pos}}
\newcommand{\wsj}{\acronym{wsj}}

\flushbottom

\title{Improved Prediction of Hospital Length of Stay for Severe Injury}
\author{Tianyu Pu \texttt{tianyu.pu@sydney.edu.au}}

\begin{document}

\makeheader

\begin{multicols}{3}

% =============================================================================
\section{Heading RGB 0, 152, 219 -- Arial 28}
\subsection{Subheading black -- Arial 24}
\noindent This is some text here...

\begin{itemize}
  \item Parsing is the process of deriving the grammatical structure for a given sentence
  \item Natural language parsing is currently slow, with the best algorithms running in $O(n^3)$ with respect to the length of the sentence
  \item There is an inherent \emph{redundancy} in natural languages where certain common phrases (or \ngrams) appear frequently in text, each time with the same grammatical structure
  \item We explore the idea of exploiting this redundancy through the construction and \emph{memoisation} of the parse structures for these frequent \ngrams
\end{itemize}

% =============================================================================
\section{Background}

% =============================================================================
\section{Problems with Pizza}

% =============================================================================
\section{Non-constituent-forming \ngrams}
Memoisation of \ngrams which do not form constituents can be achieved through the use of {\ccg}'s \emph{composition} and \emph{type raising} combinators.

\textrm{The} is joined to \textrm{potato} before \textrm{of} can be joined with \textrm{the}. Instead, we could construct the following derivation and insert it into the pre-constructed database:

Here we use \ccg forwards composition to combine \textrm{of} and \textrm{the} into a constituent-forming analysis. This chart structure could be reused to construct a span of the original phrase in the following manner:

One situation where this technique does not work is is with \emph{prepositional phrase attachment}. The correct \ccg derivation for the phrase \textrm{on the king of England} is

If we were to use in this example the same forward composed derivation of the bigram \textrm{of the} as described earlier for the bigram \textrm{on the}, the wrong analysis would be constructed.

% =============================================================================
\section{Implementation}

% =============================================================================
\section{Results}
Empirical results for the memoisation of constituent-forming \ngrams can be seen in the table below.

\begin{itemize}
  \item Different grammar models were used over varying sized \ngrams, with the  parsing time, F-score, and coverage figures being recorded
  \item No statistically significant change was observed in the F-scores or parsing times
  \item These results backup the theoretical hypothesis that persisting only constituent-forming \ngrams would not be enough to get the one structure per \ngram idea to work in practice
\end{itemize}

% =============================================================================
\section{Conclusion}
Thorough analysis of this one structure per \ngram idea using \ccg as a grammar formalism combined with a set of empirical results has shown that this one structure per \ngram idea does not work in the general when using token based \ngrams.

% =============================================================================
\section{Future work}
\begin{itemize}
  \item Recent work has been undertaken to explore the one structure per \ngram idea using \ngrams constructed of supertags instead of tokens
  \item The initial results for this new direction are positive, showing show a 10\% parser speedup being achieved with minimal loss of F-score
\end{itemize}

% =============================================================================
\references
\bibliography{../thesis/references}

\end{multicols}
\end{document}