### 2014-09-26 poster/myposter.tex
Updated structure of poster and tested citing references

\documentclass[portrait]{usydposter}
\usepackage{xspace}

\newcommand{\acronym}[1]{\textsc{#1}\xspace}
\newcommand{\cf}[1]{\mbox{$\it{#1}$}}
\newcommand{\todo}[1]{{\color{red} #1}}

\newcommand{\ngram}{n-gram\xspace}
\newcommand{\ngrams}{{\ngram}s\xspace}
\newcommand{\candc}{C\&C\xspace}
\newcommand{\ccgbank}{CCGBank\xspace}
\newcommand{\ccg}{\acronym{ccg}}
\newcommand{\cky}{\acronym{cky}}
\newcommand{\nlp}{\acronym{nlp}}
\newcommand{\np}{\acronym{np}}
\newcommand{\pos}{\acronym{pos}}
\newcommand{\wsj}{\acronym{wsj}}

\flushbottom

\title{Improved Prediction of Hospital Length of Stay for Severe Injury}
\author{Tianyu Pu \texttt{tianyu.pu@sydney.edu.au}}

\begin{document}

\makeheader

\begin{multicols}{3}

% =============================================================================
\section{Aims}
\noindent This is some text here...
\cite{Dinh2013a}

% =============================================================================
\section{Introduction}

% =============================================================================
\section{Method}

% =============================================================================
\section{Results}

% =============================================================================
\section{Discussion}

% =============================================================================
\section{Conclusions and Future work}

% =============================================================================
\references
\bibliography{../thesis/references}

\end{multicols}
\end{document}