### 2014-10-18 thesis/method/evaluation.tex
Restructure of experimental method section #12

\chapter{Evaluation} \label{chap:evaluation}

\section{Data Sets}
\subsection{Collection}
The data set used consisted of trauma registry data from the trauma centre
at the Royal Prince Alfred Hospital, a major trauma centre in New South Wales,
Australia. It covered all adult (age 15 and over) inpatient admissions to the
trauma centre from 2007--2011.

All patients were first admitted to the trauma ward until discharged
or transferred to an appropriate unit within the hospital. A single trained
data manager recorded a variety of attributes about the admitted patient,
such as age, gender, blood pressure, mechanism of injury and body regions
that were injured. We received this data in a Microsoft Excel spreadsheet.

\subsection{Characteristics}
There were 2546 patient records in the data set we received, comprising of 74
features, one of which was the target variable \textbf{LOS48}. LOS48 was a
binary variable -- that is, it could only take two values, 0 or 1 -- with 1
indicating that the patient stayed two days or less, and 0 indicating a stay
of greater than 2 days.

\todo{Perhaps insert some more information about the data itself, such
as the various attributes and their data types? Not sure if that should go
here, in the results, or in an appendix -- it is relevant but can disturb
the flow of the main text.}

\section{Performance evaluation} % don't forget cross-validation
\label{sec:perfeval}
Here we describe the methods used in order to assess the performance of the
models constructed through single classifiers and ensemble methods described
above.

\subsection{Metrics}
There are many available and commonly used metrics that we can use when
evaluating and comparing the performance of prediction models. The ones that
we use are: \todo{maybe put these in a table}
\begin{itemize}
  \item Accuracy
  \item Specificity
  \item Sensitivity
  \item Area under the receiver-operating characteristic (ROC) curve, or AUC
\end{itemize}

Since we only have one data set, we use ten-fold stratified cross-validation
to obtain values for the above metrics. We run the ten-fold cross-validation
40 times for each classifier (single or ensemble) for each of the 7 data sets
produced from feature selection. 40 runs was necessary in order for us to
compute confidence intervals for each of the metrics, which makes use of the
Central Limit Theorem from statistics.

\subsubsection{Confidence intervals}
In order to compare our results with that of Dinh et. al. \citep{Dinh2013a},
we are especially interested in the mean cross-validated AUC of our
classifiers as well as
the upper and lower bounds of the 95\% confidence interval associated with this
mean. The intervals are constructed using this argument:

For each classifier, let the population of all of its ten-fold cross-validated
AUC scores be some random variable $X$, which follows a distribution whose
parameters (such as mean and standard deviation) we do not know.
The Central Limit Theorem states that given a random
sample of $n$ items from $X$, namely $X_1,X_2,\dots,X_n$, the sample mean
$\bar{x}$
is approximately normally distributed when $n$ is at least roughly around 30
and the $X_i$'s are independent and identically distributed.
Consequently, we can construct a $(1-\alpha)$\% confidence interval for
$\bar{x}$ -- where $\alpha$ is commonly called the
\textit{significance level} -- as follows:
\begin{equation}
  \left[\bar{x} - z_{1-\alpha/2}\dfrac{s}{\sqrt{n}},
    \bar{x} + z_{1-\alpha/2}\dfrac{s}{\sqrt{n}}\right]
\end{equation}
where $\bar{x}$ is the sample mean, $\alpha$ is the significance level
(5\% or $0.05$ in our case), $s$ is the sample standard deviation, $n$
is the number of observations in our sample, and $z_{1-\alpha/2}$ is the value
of the standard normal variable $Z \sim \mathcal{N}(0,1)$ with cumulative
probability $1-\alpha/2$. Since we have taken 40 independent samples of the
ten-fold cross-validated AUC, the Central Limit Theorem holds and we can
compute the mean and standard deviation of our 40 samples in order to
construct a 95\% confidence interval for the mean AUC of each classifier.

\subsection{Statistical tests}
To decide whether or not the difference in values of each metric between
classifiers was statistically significant, paired $t$-tests were conducted for
a subset of the results: \todo{elaborate}
\begin{itemize}
  \item Some feature set vs other feature set
  \item Single classifier vs other single classifier
  \item Ensemble vs single
  \item Our best method vs Dinh2013's method
\end{itemize}