### 2014-10-24 thesis/method/preprocess.tex
Finished preprocessing section, restructure of overall methods section

\chapter{Feature Preprocessing} \label{chap:preprocess}

In the previous chapter we began by describing each training example as a
vector of feature values, and then described how a range of different
learning algorithms used those feature vectors to learn a way to
distinguish examples of one class from another. However, it is uncommon that
data can be used in a learning algorithm without first being preprocessed.
Preprocessing involves cleaning and transforming the data, and is critical
to the performance of learning algorithms that are applied to it. Without
preprocessing, it will be difficult to evaluate the true performance of
our classifiers on the data set and to draw meaningful conclusions from
comparisons between classifiers.

In this chapter give an overview of preprocessing and describe the steps we
took on our data.

\section{Cleaning} % missing values, irrelevant features
In any data mining task, adequately \textit{cleaning} the data ensures the best
chance of success: this includes deciding how to deal with missing and
inaccurate values, as well as appropriately transforming the data
\citep{Witten2005}.
The exact steps involved in data cleaning will vary between data
sets, but the overall process is similar. Common issues to deal with include:
\begin{itemize}
\item Incorrect spelling or punctuation, perhaps as a result of mistakes in
data entry
\item Leading or trailing whitespace, and other garbage characters,
perhaps as a result of data entry or
the system that stores the data
\item Missing values for various features or for the class label
\item Values for features that do not make sense (such as a patient age of
-1)
\end{itemize}
The following paragraphs address how we dealt with the issues that were
applicable to our data.

\paragraph{Incorrect spelling and punctuation.}
Since the data was initially in Excel spreadsheet format, we explored the data
by hand using Microsoft Excel's filter on each column. This allowed us to see
the range of possible values for each feature and easily detect anomalous
selling and punctuation:
\begin{itemize}
  \item In the \texttt{sex} attribute, there were values of `Female' and
  `Femal' which were consolidated into `Female'.
  \item In the \texttt{mechanism} attribute, which was textual and indicated
  the category of how the patient was injured, there were such categories as
  `Other Vehi' and `Other Vehicle' as well as `Pedal Cycl' and `Pedal Cyclist'.
  Such categories were combined into `Other Vehicle' and `Pedal Cyclist'
  respectively.
  \item Some categories in \texttt{mechanism} were also divided into whether or
  not the mechanism of injury was self-inflicted, such as `Shooting' and
  `Shooting-selfinflicted'. At the suggestion of a domain expert, we combined
  them into one category: in this case, simply
  `Shooting'; this was done for all the categories in this attribute.
  \item \texttt{disp from ED} indicated the section of the hospital that the
  patient was discharged to. This was also a textual categorical attribute.
  There were redundancies like `G. ICU' and `G.ICU' and even `GICU', which we
  all combined into one category, `GICU'.
\end{itemize}

\paragraph{Extraneous whitespace or other characters.}
After correcting for spelling and punctuation errors,
the data was then exported into CSV (comma-separated value) format, so that
we could continue to process it programmatically rather than manually without
having to use Excel. Many
fields in the data set contained leading or trailing whitespace, which was
removed with a few lines of Python code (see appendix \todo{[add reference
to appendix]}). All attribute names were converted to lowercase. In the case
where attribute names consisted of several words, the spaces were replaced
with underscores (\_) to ensure that the full attribute name was read.
For example, \texttt{Any Cancer} was converted into \texttt{any\_cancer}.

\paragraph{Missing values.}
There has been a varied treatment of missing values in the work of other
researchers, from discarding examples with missing values for any feature
to \textit{imputing} or ``filling in'' the missing values. We remove all
examples with missing feature values from consideration in all further
investigations, in line with the work of other research in LOS prediction
in the same domain \cite{Dinh2013a}.

\section{Normalisation}
Features represent different pieces of information about a training example.
Therefore, it is not surprising that not all features will have the same range
of values: a patient can have an \texttt{age} feature with a value of 20, but
their \texttt{heart rate} feature cannot possibly be 20. Learning algorithms
such as $k$-nearest neighbour (Chapter \ref{chap:classification})
are sensitive to the scales of feature values -- features with a larger range
of values will dominate a distance calculation. Therefore, we
\textit{normalise} all numeric (that is, continuous-valued)
feature values into the $[-1,1]$ range so that they are all expressed on the
same scale. Typically, values are normalised in the $[0,1]$ or the $[-1,1]$
ranges.
The formula to do this, also discussed in Chapter \ref{chap:classification},
is:
\begin{equation*}
a_i = 2\left[\dfrac{v_i - \mathrm{min }v_i}{\mathrm{max }v_i - \mathrm{min }v_i}\right] - 1
\end{equation*}
where we have used $s=2$ and $t=-1$ to transform values into the $[-1,1]$
range. This transformation was applied to each value of all numeric features,
and was performed after all cleaning steps.

\section{Discretisation}
Not all classification algorithms can handle numeric features, and some of the
ones that do require assumptions to be made about the underlying distribution
of the feature values that is not always satisfied -- recall the normality
assumption of the Na\"{i}ve Bayes classifier
from Chapter \ref{chap:classification}. This requires us
to \textit{discretise} the numeric features, which is a process of splitting
the continuous data into a small number of discrete, distinct ranges.
Additionally, there has been no work examining the use of discretisation to
improve the performance of LOS prediction models, so our work provides a
valuable guideline on what other researchers looking at a similar problem can
expect.

Two basic approaches to the discretisation problem exist: unsupervised and
supervised. \textit{Unsupervised} discretisation involves splitting each
feature without considering any class information, whereas \textit{supervised}
discretisation accounts for the class when discretising.
Unsupervised discretisation uses simple methods such as splitting the data into
fixed-width intervals -- each interval spans the same width of values,
but can contain different numbers of feature values depending on the
distribution of underlying data -- which
can often partition the continuous feature into unhelpful categories due to its
ignorance of class labels. Supervised discretisation, on the
other hand, is commonly performed using the entropy minimisation heuristic
described by Fayyad and Irani \cite{Fayyad1993}. Since we have labelled
data, we are able to use both approaches. However, we will focus on evaluating
the effect of supervised discretisation, as it is generally preferred to
unsupervised discretisation when class information about the training examples
is known. And the method proposed by Fayyad and Irani is one of the best known
general supervised discretisation techniques \cite{Witten2005}.

\subsection{Entropy-based Discretisation}
In Chapter \ref{chap:classification} we outlined the induction of a decision
tree on a data set, using the information gain as the criterion for deciding
the attribute to split on at each step. A similar approach is used in Fayyad
and Irani's entropy-based method of discretisation \cite{Fayyad1993}:
to discretise an attribute,
each possible split point is considered and the information gain from that
split is computed. The split point with the highest information gain is
therefore chosen. Once the first split is determined, the algorithm continues
recursively in the newly created upper and lower ranges.

The \textit{minimum description length (MDL)} principle is used to determine
the stopping criterion. The MDL value is calculated by adding the information
values from two cases: split or do not split. If we split, then we can encode
the split point in $\mathrm{log}_2(N-1)$ bits (where $N$ is the number of
instances), then the classes of those above and below it. If we do not split,
then each instance's label must be encoded. Mathematically, given $N$ examples
with entropy $E$, sub-intervals $k_1$ and $k_2$ with the entropy of examples
in each sub-interval denoted $E_1$ and $E_2$:
\begin{equation*}
\mathrm{MDL} = \dfrac{\mathrm{log}_2(N-1)}{N} + \dfrac{\mathrm{log}_2(3^k-2) - kE + k_1E_1 + k_2E_2}{N}
\end{equation*}
The first term is the information needed for the splitting point, and the
second is a correction necessary for encoding which classes belong to which
sub-interval. The decision rule is then to split if the information gain for
that split exceeds the MDL value: $\mathrm{gain} > \mathrm{MDL}$.

Since this requires exhaustive searching of the full range of feature values
to find potential split points, this method of supervised discretisation can
be time-consuming given large enough data sets. However, it has been proven
that any potential split that minimises the information value will never occur
between two examples of the same class. This means that we only need to
consider splits between regions of different classes, reducing the number of
possible points to consider.

\section{Conclusion}
In this chapter we discussed various issues encountered during the
preprocessing stage and how these were addressed. These steps are critical to
the success of our learning algorithms: without cleaned data, any subsequent
work is meaningless and difficult to draw valid conclusions from. By outlining
our steps, we hope to provide an example of how one can conduct the cleaning
process.

We then discussed normalisation, an important transformation of
continuous-valued data that ensures all features take on the same range of
values. This is important for ensuring that classifiers sensitive to the
relative range of values between features give valid outputs.

We concluded with outlining discretisation, in particular Fayyad and Irani's
entropy-based method. Such a technique has not been used in LOS prediction
and by including it in our work, we can evaluate its effectiveness in this
domain and let our results serve
as a baseline and reference for future work.