### 2014-10-25 thesis/conclusion/conclusion.tex
Update structure of conclusions

\chapter{Conclusions and Future Work} \label{chap:conclusion}

\section{Conclusions}

\section{Future Work}