### 2014-10-28 thesis/appendix/allresults.tex
Add appendices and results for trauma data

\chapter{Full Results Listing} \label{app:results}

\section{Trauma LOS}
\subsection{Performance Results}

\input{appendix/tr-all-summary.tex}

\input{appendix/tr-dinh-summary.tex}

\input{appendix/tr-expert-summary.tex}

\input{appendix/tr-cfs-summary.tex}

\input{appendix/tr-wrapper-summary.tex}

\input{appendix/tr-corr-threshold-disc.tex}
\input{appendix/tr-corr-threshold-nodisc.tex}

\input{appendix/tr-ig-threshold-summary-disc.tex}
\input{appendix/tr-ig-threshold-summary-nodisc.tex}

\input{appendix/tr-oner-threshold-summary-disc.tex}
\input{appendix/tr-oner-threshold-summary-nodisc.tex}

\subsection{Feature Selection Results}

\input{appendix/tr-nothreshold-features.tex}

\input{appendix/tr-corr-threshold-features.tex}

\input{appendix/tr-ig-threshold-features.tex}

\input{appendix/tr-oner-threshold-features.tex}

\section{General LOS}
\subsection{Performance Results}

\subsection{Feature Selection Results}