### 2014-10-28 thesis/appendix/paper.tex

\chapter{Artificial Intelligence in Medicine -- Submitted Paper} \label{app:paper}