### 2014-10-28 thesis/appendix/tr-corr-threshold-features.tex

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Pearson Correlation - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 57 & 73.08\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.01 & 54 & 69.23\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.05 & 41 & 52.56\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, faceany, any\_heart, any\_cancer, face3, upperlimboponly \\ \hline
0.1 & 29 & 37.18\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor, mentalhealth, rr, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface \\ \hline
0.125 & 20 & 25.64\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson, gcs, triss, bp, age, chest3, anypelvis, head3, headminor \\ \hline
0.15 & 12 & 15.38\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3, bodyregions, operation, lowerlimbnopelvis, neurosurgery, charlson \\ \hline
0.2 & 7 & 8.97\% & traumatier, icu, iss, upperlimbonly, lower\_limbop, lowerlimbany, lower3 \\ \hline
0.3 & 1 & 1.28\% & traumatier \\ \hline
0.5 & 0 & 0.00\% & - \\ \hline
0.75 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{11cm}|}
\hline
\multicolumn{ 4}{|c|}{Pearson Correlation - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 58 & 74.36\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.01 & 55 & 70.51\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.05 & 41 & 52.56\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.1 & 27 & 34.62\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.125 & 19 & 24.36\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery, chest3, anypelvis, headminor, head3, mentalhealth, normalvitals, arrivalmode, spine3, headany, laparotomy, spineany, headface, chestany, rr, abdo3, penetrating, abdoany, any\_stroketia, mechcode, any\_dm, pr, faceany, any\_heart, any\_cancer, face3, upperlimboponly, any\_liver, any\_kidney, thoracotomy, bp, married, upperlimbany, pelvic\_fracture, ext3, alcohol, sex, neckany, upper3, transfer, neck3 \\ \hline
0.15 & 15 & 19.23\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss, charlson, age, gcs, operation, lowerlimbnopelvis, neurosurgery \\ \hline
0.2 & 9 & 11.54\% & iss, traumatier, icu, upperlimbonly, bodyregions, lower\_limbop, lowerlimbany, lower3, triss \\ \hline
0.3 & 2 & 2.56\% & iss, traumatier \\ \hline
0.5 & 0 & 0.00\% & - \\ \hline
0.75 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}