### 2014-10-29 thesis/appendix/pt-nothreshold-features.tex

\begin{table}[h]
\caption{}
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{No Discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
C4.5 Wrapper & 4 & 28.57\% & Education, Episode Type, Inpatient Service, Main Diagnosis \\ \hline
\end{tabular}
\vspace{1cm}
\vfill
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
C4.5 Wrapper & 4 & 28.57\% & Education, Episode Type, Inpatient Service, Main Diagnosis \\ \hline
\end{tabular}
\end{table}