### 2014-10-29 thesis/appendix/pt-oner-threshold-features.tex

\begin{table}[htbp]
\caption{}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
67.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
68.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
69.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
71.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
74.405 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
76.905 & 6 & 42.86\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure \\ \hline
81.905 & 3 & 21.43\% & EpisodeType, InpatientService, OriginEpisodeType \\ \hline
86.905 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
96.905 & 0 & 0.00\% &  \\ \hline
\end{tabular}}
\end{table}

% nodisc is exactly the same