### 2014-10-30 thesis/appendix/tr-ig-threshold-features.tex
Complete all results tables for the trauma data set

\begin{table}[htbp]
\caption{Features selected from the discretised trauma data set at varying thresholds of information gain. Features shown are in order from highest to lowest information gain.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Information Gain - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 57 & 73.08\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.001 & 47 & 60.26\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany \\ \hline
0.005 & 33 & 42.31\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating \\ \hline
0.01 & 27 & 34.62\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, rr, mentalhealth, spine3, normalvitals \\ \hline
0.05 & 4 & 5.13\% & iss, triss, traumatier, icu \\ \hline
0.1 & 2 & 2.56\% & iss, triss \\ \hline
0.15 & 1 & 1.28\% & iss \\ \hline
0.2 & 0 & 0.00\% & - \\ \hline
0.3 & 0 & 0.00\% & - \\ \hline
0.5 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}

\begin{table}[htbp]
\caption{Features selected from the non-discretised trauma data set at varying thresholds of information gain. Features shown are in order from highest to lowest information gain.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Information Gain - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
0 & 57 & 73.08\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany, ext3, alcohol, sex, neckany, upper3, transfer, neck3, upper\_limbop, externalany, english \\ \hline
0.001 & 47 & 60.26\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating, abdoany, any\_stroketia, any\_cancer, any\_dm, any\_heart, faceany, face3, upperlimboponly, any\_kidney, thoracotomy, any\_liver, married, pelvic\_fracture, upperlimbany \\ \hline
0.005 & 33 & 42.31\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals, spineany, headany, headface, chestany, abdo3, penetrating \\ \hline
0.01 & 27 & 34.62\% & iss, triss, traumatier, icu, bodyregions, arrivalmode, lower\_limbop, lower3, upperlimbonly, lowerlimbany, charlson, age, gcs, mechcode, neurosurgery, operation, rr, lowerlimbnopelvis, bp, anypelvis, chest3, head3, headminor, laparotomy, mentalhealth, spine3, normalvitals \\ \hline
0.05 & 5 & 6.41\% & iss, triss, traumatier, icu, bodyregions \\ \hline
0.1 & 2 & 2.56\% & iss, triss \\ \hline
0.15 & 1 & 1.28\% & iss \\ \hline
0.2 & 0 & 0.00\% & -  \\ \hline
0.3 & 0 & 0.00\% & - \\ \hline
0.5 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}