### 2014-10-30 thesis/appendix/tr-oner-threshold-features.tex

\begin{table}[htbp]
\caption{Features selected from the trauma data set using varying accuracy thresholds for the 1R feature selector. Features are shown in ranked order, from highest to lowest accuracy.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 59 & 75.64\% & iss, upperlimbonly, triss, arrivalmode, mechcode, ext3, chestany, headface, faceany, neckany, any\_stroketia, headany, bodyregions, externalany, lowerlimbany, spineany, upperlimbany, abdoany, traumateam, penetrating, face3, normalvitals, bp, sex, pr, rr, gcs, traumatier, transfer, icu, head3, chest3, neck3, headminor, operation, alcohol, mentalhealth, upperlimboponly, lowerlimbnopelvis, english, any\_heart, any\_cancer, any\_kidney, any\_liver, charlson, married, any\_dm, laparotomy, upper3, spine3, abdo3, lower3, thoracotomy, lower\_limbop, neurosurgery, upper\_limbop, pelvic\_fracture, anypelvis, age \\ \hline
67.905 & 4 & 5.13\% & iss, upperlimbonly, triss, arrivalmode \\ \hline
68.905 & 2 & 2.56\% & iss, upperlimbonly \\ \hline
69.905 & 2 & 2.56\% & iss, upperlimbonly \\ \hline
71.905 & 0 & 0.00\% & - \\ \hline
74.405 & 0 & 0.00\% & - \\ \hline
76.905 & 0 & 0.00\% & - \\ \hline
81.905 & 0 & 0.00\% & - \\ \hline
86.905 & 0 & 0.00\% & - \\ \hline
96.905 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\vspace{1cm}
\vfill
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - No discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 55 & 70.51\% & triss, iss, upperlimbonly, arrivalmode, mechcode, bodyregions, ext3, chestany, spineany, abdoany, head3, upperlimbany, lowerlimbany, headface, externalany, neckany, any\_stroketia, faceany, headany, gcs, normalvitals, sex, icu, transfer, traumatier, neck3, penetrating, traumateam, face3, chest3, any\_dm, headminor, operation, alcohol, mentalhealth, upperlimboponly, lowerlimbnopelvis, english, any\_heart, any\_cancer, any\_kidney, any\_liver, abdo3, charlson, married, thoracotomy, lower3, upper3, spine3, lower\_limbop, laparotomy, neurosurgery, anypelvis, upper\_limbop, pelvic\_fracture \\ \hline
67.905 & 4 & 5.13\% & triss, iss, upperlimbonly, arrivalmode \\ \hline
68.905 & 3 & 3.85\% & triss, iss, upperlimbonly \\ \hline
69.905 & 3 & 3.85\% & triss, iss, upperlimbonly \\ \hline
71.905 & 1 & 1.28\% & triss \\ \hline
74.405 & 0 & 0.00\% & - \\ \hline
76.905 & 0 & 0.00\% & - \\ \hline
81.905 & 0 & 0.00\% & - \\ \hline
86.905 & 0 & 0.00\% & - \\ \hline
96.905 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}