### 2014-10-30 thesis/appendix/pt-nothreshold-features.tex

\begin{table}[htbp]
\caption{Features selected using automatic methods which select the best subset and do not require a threshold (CFS and C4.5 wrapper). Features are not ranked with these feature selectors. The same features are selected by each method regardless of discretisation or not.}
\begin{tabular}{|l|r|r|p{8.5cm}|}
\hline
\multicolumn{ 4}{|c|}{Discretisation/No Discretisation} \\ \hline
Method & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
CFS & 2 & 14.29\% & Episode Type, Main Diagnosis \\ \hline
C4.5 Wrapper & 4 & 28.57\% & Education, EpisodeType, InpatientService, MainDiagnosis \\ \hline
\end{tabular}
\end{table}