### 2014-10-30 thesis/appendix/pt-oner-threshold-features.tex

\begin{table}[htbp]
\caption{Features selected from the general hospital data set using varying accuracy thresholds for the 1R feature selector. Features are shown in ranked order, from highest to lowest accuracy. The same features are selected regardless of discretisation.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|r|r|r|p{9.5cm}|}
\hline
\multicolumn{ 4}{|c|}{1R Accuracy - Discretisation/No Discretisation} \\ \hline
\multicolumn{1}{|l|}{Threshold} & \multicolumn{1}{l|}{Num selected} & \multicolumn{1}{l|}{\% of total} & Features \\ \hline
66.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
67.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
68.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
69.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
71.905 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
74.405 & 14 & 100.00\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure, AdmissionHour, Age, MaritalStatus, Education, prev\_adm, AdmissionDay, AdmissionMonth, Sex \\ \hline
76.905 & 6 & 42.86\% & EpisodeType, InpatientService, OriginEpisodeType, MedicalSpecialty, MainDiagnosis, MainProcedure \\ \hline
81.905 & 3 & 21.43\% & EpisodeType, InpatientService, OriginEpisodeType \\ \hline
86.905 & 2 & 14.29\% & EpisodeType, InpatientService \\ \hline
96.905 & 0 & 0.00\% & - \\ \hline
\end{tabular}}
\end{table}

% nodisc is exactly the same