### 2014-10-30 thesis/appendix/pt-rankeddist-summary.tex

\begin{sidewaystable}[htbp]
\vspace{15.7cm}
\caption{Results for the $k$-NN classifier with our proposed Ranked Distance approach on the general hospital data set with and without discretisation, for varying numbers of numbers $k$ and decay constant $c$.}
\medskip
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|l|*{10}{cc|}}
\hline
 & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
 & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
k & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- \\ \hline
  1 & 92.77\% & 0.04\% & 96.45\% & 0.05\% & 94.72\% & 0.02\% & 96.86\% & 0.03\% & 94.45\% & 0.04\% & 96.71\% & 0.03\% & 94.62\% & 0.02\% & 96.71\% & 0.03\% & 94.38\% & 0.03\% & 96.51\% & 0.03\% \\ \hline
  5 & 93.61\% & 0.06\% & 96.64\% & 0.03\% & 95.27\% & 0.03\% & 97.05\% & 0.04\% & 95.08\% & 0.03\% & 96.96\% & 0.02\% & 95.15\% & 0.03\% & 96.98\% & 0.04\% & 94.96\% & 0.03\% & 96.77\% & 0.02\% \\ \hline
  10 & 93.76\% & 0.05\% & 96.92\% & 0.04\% & 95.38\% & 0.04\% & 97.25\% & 0.04\% & 95.04\% & 0.04\% & 97.18\% & 0.02\% & 95.10\% & 0.04\% & 97.17\% & 0.02\% & 94.97\% & 0.03\% & 97.02\% & 0.03\% \\ \hline
  15 & 94.42\% & 0.04\% & 97.15\% & 0.03\% & 95.57\% & 0.02\% & 97.43\% & 0.02\% & 95.38\% & 0.04\% & 97.35\% & 0.02\% & 95.43\% & 0.03\% & 97.29\% & 0.02\% & 95.17\% & 0.02\% & 97.17\% & 0.02\% \\ \hline
  20 & 95.69\% & 0.03\% & 97.33\% & 0.03\% & 96.30\% & 0.03\% & 97.56\% & 0.03\% & 96.14\% & 0.03\% & 97.47\% & 0.03\% & 96.21\% & 0.03\% & 97.41\% & 0.02\% & 95.97\% & 0.02\% & 97.27\% & 0.02\% \\ \hline
\end{tabular}}
\label{}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|l|*{10}{cc|}}
\hline
 & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
 & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
k & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- \\ \hline
  1 & 0.904 & 0.0009 & 0.957 & 0.0008 & 0.963 & 0.0005 & 0.984 & 0.0003 & 0.970 & 0.0005 & 0.988 & 0.0004 & 0.973 & 0.0004 & 0.990 & 0.0003 & 0.974 & 0.0003 & 0.990 & 0.0002 \\ \hline
  5 & 0.917 & 0.0009 & 0.960 & 0.0006 & 0.969 & 0.0007 & 0.985 & 0.0003 & 0.976 & 0.0003 & 0.989 & 0.0004 & 0.978 & 0.0005 & 0.991 & 0.0000 & 0.978 & 0.0002 & 0.991 & 0.0003 \\ \hline
  10 & 0.920 & 0.0009 & 0.964 & 0.0008 & 0.969 & 0.0004 & 0.987 & 0.0003 & 0.976 & 0.0005 & 0.991 & 0.0004 & 0.977 & 0.0004 & 0.992 & 0.0002 & 0.979 & 0.0000 & 0.992 & 0.0003 \\ \hline
  15 & 0.929 & 0.0009 & 0.966 & 0.0005 & 0.972 & 0.0004 & 0.988 & 0.0003 & 0.977 & 0.0002 & 0.991 & 0.0002 & 0.979 & 0.0004 & 0.992 & 0.0000 & 0.980 & 0.0004 & 0.993 & 0.0003 \\ \hline
  20 & 0.947 & 0.0006 & 0.969 & 0.0004 & 0.979 & 0.0003 & 0.988 & 0.0004 & 0.984 & 0.0004 & 0.991 & 0.0003 & 0.985 & 0.0003 & 0.992 & 0.0004 & 0.986 & 0.0000 & 0.993 & 0.0002 \\ \hline
\end{tabular}}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|l|*{10}{cc|}}
\hline
 & \multicolumn{20}{c|}{Decay constant $c$ - No Discretisation} \\ \hline
 & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
k & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- & Acc & +/- \\ \hline
  1 & 92.89\% & 0.04\% & 96.48\% & 0.04\% & 94.81\% & 0.04\% & 97.07\% & 0.02\% & 94.52\% & 0.01\% & 96.83\% & 0.03\% & 94.69\% & 0.03\% & 96.83\% & 0.02\% & 94.41\% & 0.02\% & 96.61\% & 0.01\% \\ \hline
  5 & 93.70\% & 0.05\% & 96.68\% & 0.02\% & 95.44\% & 0.04\% & 97.21\% & 0.03\% & 95.10\% & 0.04\% & 97.10\% & 0.03\% & 95.20\% & 0.04\% & 97.10\% & 0.02\% & 95.08\% & 0.04\% & 96.87\% & 0.03\% \\ \hline
  10 & 93.93\% & 0.05\% & 96.94\% & 0.02\% & 95.56\% & 0.03\% & 97.42\% & 0.03\% & 95.30\% & 0.03\% & 97.26\% & 0.02\% & 95.35\% & 0.04\% & 97.27\% & 0.02\% & 95.16\% & 0.03\% & 97.10\% & 0.02\% \\ \hline
  15 & 94.56\% & 0.04\% & 97.20\% & 0.03\% & 95.90\% & 0.03\% & 97.55\% & 0.02\% & 95.60\% & 0.02\% & 97.41\% & 0.03\% & 95.68\% & 0.02\% & 97.37\% & 0.02\% & 95.40\% & 0.02\% & 97.24\% & 0.01\% \\ \hline
  20 & 95.75\% & 0.02\% & 97.38\% & 0.03\% & 96.52\% & 0.03\% & 97.63\% & 0.01\% & 96.30\% & 0.04\% & 97.52\% & 0.01\% & 96.36\% & 0.01\% & 97.49\% & 0.03\% & 96.13\% & 0.02\% & 97.32\% & 0.02\% \\ \hline
\end{tabular}}
\vspace{0.5cm}
\vfill
\resizebox{\linewidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{|l|*{10}{cc|}}
\hline
 & \multicolumn{20}{c|}{Decay constant $c$ - Discretisation} \\ \hline
 & \multicolumn{ 2}{c|}{0.1} & \multicolumn{ 2}{c|}{0.2} & \multicolumn{ 2}{c|}{0.3} & \multicolumn{ 2}{c|}{0.4} & \multicolumn{ 2}{c|}{0.5} & \multicolumn{ 2}{c|}{0.6} & \multicolumn{ 2}{c|}{0.7} & \multicolumn{ 2}{c|}{0.8} & \multicolumn{ 2}{c|}{0.9} & \multicolumn{ 2}{c|}{1} \\
k & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- & AUC & +/- \\ \hline
  1 & 0.905 & 0.0011 & 0.956 & 0.0007 & 0.965 & 0.0007 & 0.984 & 0.0003 & 0.973 & 0.0004 & 0.989 & 0.0003 & 0.975 & 0.0003 & 0.990 & 0.0003 & 0.977 & 0.0002 & 0.991 & 0.0003 \\ \hline
  5 & 0.917 & 0.0012 & 0.958 & 0.0006 & 0.969 & 0.0006 & 0.986 & 0.0005 & 0.977 & 0.0004 & 0.990 & 0.0004 & 0.979 & 0.0004 & 0.991 & 0.0003 & 0.980 & 0.0003 & 0.992 & 0.0000 \\ \hline
  10 & 0.922 & 0.0006 & 0.962 & 0.0006 & 0.971 & 0.0005 & 0.988 & 0.0002 & 0.978 & 0.0003 & 0.991 & 0.0003 & 0.980 & 0.0003 & 0.993 & 0.0003 & 0.981 & 0.0003 & 0.993 & 0.0003 \\ \hline
  15 & 0.930 & 0.0011 & 0.966 & 0.0003 & 0.974 & 0.0004 & 0.988 & 0.0004 & 0.979 & 0.0004 & 0.991 & 0.0003 & 0.982 & 0.0003 & 0.993 & 0.0003 & 0.982 & 0.0003 & 0.993 & 0.0000 \\ \hline
  20 & 0.946 & 0.0008 & 0.967 & 0.0007 & 0.979 & 0.0004 & 0.988 & 0.0003 & 0.986 & 0.0003 & 0.992 & 0.0003 & 0.987 & 0.0002 & 0.993 & 0.0003 & 0.987 & 0.0004 & 0.993 & 0.0002 \\ \hline
\end{tabular}}
\end{sidewaystable}