### 2014-10-31 thesis/results/overall-trauma-auc.tex

\begin{table}[htbp]
\caption{Overall AUC performance of each classifier with a particular feature selection method on the trauma data set. `D' indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied.}
\resizebox{\textwidth}{!}{%
\begin{tabular}{|l|cccccccc|}
\hline
Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 & 0.498 \\
1R & 0.67 & 0.5 & \begin{tabular}[c]{@{}c@{}}0.581\\ D\end{tabular} & 0.668 & 0.668 & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.005\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.67\\ t=69.905\end{tabular}} \\
NB & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & 0.804 & \begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.815\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.829\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.8\\ D, t=66.905\end{tabular} \\
DT & \begin{tabular}[c]{@{}c@{}}0.756\\ D\end{tabular} & 0.772 & \begin{tabular}[c]{@{}c@{}}0.786\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.793\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.806\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.766\\ D, t=66.905\end{tabular} \\
LR & \begin{tabular}[c]{@{}c@{}}0.844\\ D\end{tabular} & 0.812 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.819\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.846\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.811\\ D, t=66.905\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}0.843\\ D\end{tabular} & 0.787 & 0.807 & \begin{tabular}[c]{@{}c@{}}0.782\\ D\end{tabular} & 0.806 & \textbf{\begin{tabular}[c]{@{}c@{}}0.844\\ D, t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.724\\ D, t=66.905\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}0.726\\ D\end{tabular} & 0.719 & \begin{tabular}[c]{@{}c@{}}0.760\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.798\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.816\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798\\ D, t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.778\\ t=68.905\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}0.827\\ D\end{tabular} & 0.792 & \begin{tabular}[c]{@{}c@{}}0.808\\ D\end{tabular} & 0.814 & \begin{tabular}[c]{@{}c@{}}0.828\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.831\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.832\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ D, t=66.905\end{tabular} \\
RD & \begin{tabular}[c]{@{}c@{}}0.834\\ D\end{tabular} & 0.742 & \begin{tabular}[c]{@{}c@{}}0.807\\ D\end{tabular} & 0.81 & \begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.836\\ D, t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834\\ D, t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.791\\ D, t=66.905\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}0.768\\ D\end{tabular} & 0.774 & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.815 & \textbf{\begin{tabular}[c]{@{}c@{}}0.838\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.82\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}0.803\\ D\end{tabular} & 0.781 & \begin{tabular}[c]{@{}c@{}}0.784\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.825\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ D, t=68.905\end{tabular} \\ \hline
\end{tabular}
}
\end{table}