### 2014-10-31 thesis/results/results.tex

\chapter{Results and Discussion} \label{chap:results}

\section{Overall Performance}
\subsection{Trauma Data Set}
\input{results/overall-trauma-acc.tex}

\input{results/overall-trauma-auc.tex}

\subsection{General Data Set}

\section{Comparison of Classifiers}

\section{Comparison of Feature Selection Methods}

\section{Effect of Discretisation}