### 2014-11-01 thesis/results/features-tr-auc.tex

\begin{table}[htbp]
\caption{Comparison of AUC between all feature selection methods for each classifier with and without discretisation, on the trauma data set. The best AUC for each feature selection method is highlighted in bold, and the best AUC for each classifier is italicised.}
\label{tab:features-tr-auc}
\resizebox{!}{11cm}{%
\begin{tabular}{|l|cccccccc|}
 \hline
 & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
\textit{1R} &  &  &  &  &  &  &  &  \\
Discretised & 0.569 & 0.5 & 0.581 & 0.569 & 0.581 & \begin{tabular}[c]{@{}c@{}}0.572\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.581\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.571\\ t=68.905\end{tabular} \\
Not discretised & 0.67 & 0.5 & 0.581 & 0.668 & 0.668 & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.669\\ t=0.005\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.67\\ t=69.905\end{tabular}} \\
\textit{NB} &  &  &  &  &  &  &  &  \\
Discretised & 0.828 & 0.804 & 0.806 & 0.815 & 0.827 & \begin{tabular}[c]{@{}c@{}}0.829\\ t=0.125\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.832\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.8\\ t=66.905\end{tabular} \\
Not discretised & 0.812 & 0.804 & 0.793 & 0.801 & 0.786 & \begin{tabular}[c]{@{}c@{}}0.815\\ t=0.15\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.817\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.784\\ t=66.905\end{tabular} \\
\textit{DT} &  &  &  &  &  &  &  &  \\
Discretised & 0.756 & 0.772 & 0.786 & 0.793 & \textit{0.806} & \begin{tabular}[c]{@{}c@{}}0.799\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.766\\ t=66.905\end{tabular} \\
Not discretised & 0.727 & 0.772 & 0.762 & 0.782 & 0.784 & \textit{\begin{tabular}[c]{@{}c@{}}0.783\\ t=0.2\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.757\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.756\\ t=69.905\end{tabular} \\
\textit{LR} &  &  &  &  &  &  &  &  \\
Discretised & \textbf{0.844} & \textbf{0.812} & \textbf{0.825} & \textbf{0.819} & 0.831 & \textit{\textbf{\begin{tabular}[c]{@{}c@{}}0.846\\ t=0.1\end{tabular}}} & \textbf{\begin{tabular}[c]{@{}c@{}}0.844\\ t=0\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}0.811\\ t=66.905\end{tabular}} \\
Not discretised & \textit{0.837} & \textbf{0.812} & 0.812 & 0.812 & 0.809 & \textit{\begin{tabular}[c]{@{}c@{}}0.837\\ t=0\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.837\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.798\\ t=66.905\end{tabular} \\
\textit{SVM} &  &  &  &  &  &  &  &  \\
Discretised & 0.843 & 0.787 & 0.767 & 0.782 & 0.796 & \textit{\begin{tabular}[c]{@{}c@{}}0.844\\ t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.842\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.724\\ t=66.905\end{tabular} \\
Not discretised & \textit{0.836} & 0.787 & 0.807 & 0.778 & 0.806 & \textit{\begin{tabular}[c]{@{}c@{}}0.836\\ t=0\end{tabular}} & \textit{\begin{tabular}[c]{@{}c@{}}0.836\\ t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.704\\ t=66.905\end{tabular} \\
\textit{1NN} &  &  &  &  &  &  &  &  \\
Discretised & 0.726 & 0.719 & 0.76 & 0.798 & \textit{0.816} & \begin{tabular}[c]{@{}c@{}}0.798\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.763\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.771\\ t=68.905\end{tabular} \\
Not discretised & 0.651 & 0.719 & 0.653 & 0.772 & 0.756 & \begin{tabular}[c]{@{}c@{}}0.752\\ t=0.3\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.768\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.778\\ t=68.905\end{tabular}} \\
\textit{20NN} &  &  &  &  &  &  &  &  \\
Discretised & 0.827 & 0.792 & 0.808 & 0.811 & 0.828 & \begin{tabular}[c]{@{}c@{}}0.831\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.832\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.799\\ t=66.905\end{tabular} \\
Not discretised & 0.799 & 0.792 & 0.77 & \textit{0.814} & 0.803 & \begin{tabular}[c]{@{}c@{}}0.812\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.81\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
\textit{RD} &  &  &  &  &  &  &  &  \\
Discretised & 0.834 & 0.742 & 0.807 & 0.806 & 0.825 & \textit{\begin{tabular}[c]{@{}c@{}}0.836\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.834\\ t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.791\\ t=66.905\end{tabular} \\
Not discretised & 0.826 & 0.742 & 0.805 & 0.81 & 0.811 & \begin{tabular}[c]{@{}c@{}}0.83\\ t=0.1\end{tabular} & \textit{\begin{tabular}[c]{@{}c@{}}0.831\\ t=0.01\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.788\\ t=68.905\end{tabular} \\
\textit{K*} &  &  &  &  &  &  &  &  \\
Discretised & 0.768 & 0.774 & 0.803 & 0.813 & \textit{\textbf{0.838}} & \begin{tabular}[c]{@{}c@{}}0.82\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.802\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.782\\ t=67.905\end{tabular} \\
Not discretised & 0.732 & 0.774 & 0.784 & \textit{0.815} & 0.809 & \begin{tabular}[c]{@{}c@{}}0.811\\ t=0.2\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.779\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.787\\ t=68.905\end{tabular} \\
\textit{MLP} &  &  &  &  &  &  &  &  \\
Discretised & 0.803 & 0.781 & 0.784 & 0.802 & \textit{0.825} & \begin{tabular}[c]{@{}c@{}}0.807\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.803\\ t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.772\\ t=68.905\end{tabular} \\
Not discretised & 0.78 & 0.781 & 0.774 & 0.799 & 0.804 & \textit{\begin{tabular}[c]{@{}c@{}}0.818\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.788\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.752\\ t=68.905\end{tabular} \\ \hline
\end{tabular}
}
\end{table}