### 2014-11-01 thesis/results/overall-trauma-acc.tex

\begin{table}[htbp]
\label{tab:overall-tr-acc}
\caption{Overall best accuracy of each classifier with a particular feature selection method on the trauma data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\resizebox{\linewidth}{!}{%
\begin{tabular}{|l|cccccccc|}
\hline
Classifier & None & Baseline & Expert & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* & 66.91* \\
1R & 71.95 & 66.91* & \begin{tabular}[c]{@{}c@{}}70.64\\ D\end{tabular} & \textbf{72.01} & 71.93 & \begin{tabular}[c]{@{}c@{}}71.91\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.95\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.97\\ t=66.905\end{tabular} \\
NB & \begin{tabular}[c]{@{}c@{}}73.83\\ D\end{tabular} & 74.62* & \begin{tabular}[c]{@{}c@{}}73.19\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.4\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}75.8\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.85\\ D, t=0.125\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.45\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.37\\ D, t=69.905\end{tabular} \\
DT & \begin{tabular}[c]{@{}c@{}}75\\ D\end{tabular} & 74.38* & \begin{tabular}[c]{@{}c@{}}76.39\\ D\end{tabular} & 75.4 & \textbf{\begin{tabular}[c]{@{}c@{}}77.52\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.05\\ D, t=0.11\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.17\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.49\\ t=66.905\end{tabular} \\
LR & \begin{tabular}[c]{@{}c@{}}77.62\\ D\end{tabular} & 75.06* & \begin{tabular}[c]{@{}c@{}}76.45\\ D\end{tabular} & 74.14 & \begin{tabular}[c]{@{}c@{}}76.3\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.7\\ D, t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}77.65\\ D, t=0\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.4\\ D, t=66.905\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}77.39\\ D\end{tabular} & 74.47* & \begin{tabular}[c]{@{}c@{}}75.67\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}72.57\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}75.26\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}77.42\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.53\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}71.22\\ D\end{tabular} & 70.28* & \begin{tabular}[c]{@{}c@{}}72.56\\ D\end{tabular} & 74.44 & \textbf{\begin{tabular}[c]{@{}c@{}}77.81\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.75\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.78\\ t=68.905\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}75.9\\ D\end{tabular} & 73.59* & \begin{tabular}[c]{@{}c@{}}75.51\\ D\end{tabular} & 75.3 & \begin{tabular}[c]{@{}c@{}}76.48\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.26\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.56\\ D, t=0.005\end{tabular}} & \begin{tabular}[c]{@{}c@{}}73.56\\ t=68.905\end{tabular} \\
RD & \begin{tabular}[c]{@{}c@{}}76.92\\ D\end{tabular} & 69.68* & \begin{tabular}[c]{@{}c@{}}76.19\\ D\end{tabular} & 74.38 & \textbf{\begin{tabular}[c]{@{}c@{}}77.22\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.87\\ D, t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.83\\ D, t=66.905\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}71.4\\ D\end{tabular} & 72.20* & \begin{tabular}[c]{@{}c@{}}75.08\\ D\end{tabular} & 75.58 & \textbf{\begin{tabular}[c]{@{}c@{}}77.27\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08\\ D, t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.45\\ t=68.905\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}73.79\\ D\end{tabular} & 73.03* & \begin{tabular}[c]{@{}c@{}}73.45\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.38\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03\\ D, t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}71.94\\ D, t=66.905\end{tabular} \\ \hline
\end{tabular}
}
\end{table}