### 2014-11-03 thesis/conclusion/conclusion.tex

\chapter{Conclusions and Future Work} \label{chap:conclusion}

In this thesis we conducted a systematic and empirical evaluation of a number
of classification algorithms and feature selection methods, as well as the
effect of discretisation, for predicting the length of hospital stay of trauma
patients. In particular, we introduced a new classification algorithm, Ranked
Distance Nearest Neighbours, and compared its accuracy and AUC with ten
other classifiers -- Zero-R, 1R, Na\"{i}ve Bayes, C4.5 decision tree, logistic
regression, support vector machine, $k$-nearest neighbour with $k=1$ and 20,
K* and multi-layer perceptron -- several of which (for example K*) have not
been used in existing work on LOS prediction. We also investigated a variety of
feature selection methods -- CFS, C4.5 wrapper, correlation coefficient,
information gain, 1R and expert selection -- encompassing both manual and
automatic methods, many of which have not been applied to the problem of LOS
prediction. Our experiments were conducted on two data sets, one from the
trauma ward of a Sydney hospital and the other a general hospital-wide data
set from Portugal.

\section{Conclusions}
The best accuracy we achieved on the trauma data was 77.81\% with the $k$-NN
classifier using 1 nearest neighbour and 11 out of 78 features. This represents
an improvement over previous work done by Dinh et al. \cite{Dinh2013a}
on the same data set,
where they managed to achieve 75.06\% with logistic regression. More
importantly, we were able to improve upon their AUC of 0.812 with a result of
0.846, also using logistic regression but first discretising numeric features
and reducing the feature set to 29 features.
If we consider an even smaller number of features, then K*
gives an AUC of 0.838 with 11 features selected by the C4.5 wrapper, which
improves upon the result of Dinh et al. \cite{Dinh2013a} while using a little
more than half
the number of features. The drawback of $k$-NN and K* is that they require all
examples to be stored and then searched during classification, which becomes
storage- and computation-expensive as the data set increases in size.

We were able to achieve substantially better results on the hospital-wide data,
with a best accuracy of 98.23\% using a C4.5 decision tree and best AUC of
0.994 achieved equally by logistic regression, SVM and K*. Like with the
trauma data, we were also able to sacrifice a small amount of accuracy or AUC
in order to use substantially less features: the MLP was able to achieve
96.45\% accuracy and 0.983 AUC while using only 2 out of 14 features. In both
these data sets, investigating feature selection was not only a way to
improve the performance of our classifiers, but also allowed us to see the
trade-off between the two (sometimes conflicting) objectives of increasing
predictive power and reducing feature set size.

An important contribution of our work is the use of several nearest neighbour
algorithms for predicting the LOS class of a patient. We found that these
techniques can achieve good accuracy and discriminating ability if a suitable
set of features is found. In addition, we proposed a new algorithm, Ranked
Distance Nearest Neighbours, which is an extension of the standard nearest
neighbours classifier that takes the relative importance
of each feature into account when determining the nearest neighbours of an
example. We found that this achieved better results than the other nearest
neighbour algorithms we evaluated, but only when most of the original feature
set was used.

Although we were able to improve on the work of Dinh et al. \cite{Dinh2013a}
and confirm the
importance of some features they and others found in predicting the LOS for
trauma patients, we did not find a single classifier or feature selector
that was superior to the others but instead observed different combinations
giving better results than others in certain situations. This highlights the
importance of understanding the data and investigating a range of classifiers
and feature selection techniques when approaching any medical data mining
problem, for which our work will become a catalogue.
However, we did find that logistic regression tended to perform well
on both data sets and across a number of feature selection methods, supporting
its common use in deriving classifiers for various medical phenomena.

\section{Future Work}
In this thesis we have systematically evaluated a large number of combinations
of classifiers and feature selection methods while also testing the effect of
supervised discretisation, and have thus created a catalogue of examples of
the application of data mining techniques to LOS classification. Our work on
this topic has raised a number of avenues for futher research, one of which is
to re-formulate the binary classification problem as a \textit{multi-class}
classification problem. Our current categorisation of the LOS outcome is very
crude: we can only classify patients into two categories. In order to better
assist physicians in making decisions, we could split the classification into
three or more classes, such as 1-3 days, 4-10 days and $>$11 days, and
investigate how different classifiers and features perform for this slightly
different problem.

We mentioned in the discussion in Chapter \ref{chap:results} that we did not
fine-tune the parameters of the classifiers that we evaluated. There is room
for us to optimise these parameters and test whether these result in
significant improvements in accuracy and AUC, and explore reasons why this
did or did not help. We could also broaden our investigation by additionally
using another data set, preferably from a different medical domain.

In our proposed Ranked Distance Nearest Neighbours algorithm, we only considered
one kind of weight assignment function, but it is not the only way to assign
these weights. There is room to investigate other types of functions, as well
as using weights obtained from another algorithm (such as logistic regression).
In addition to assigning the weights, we have also only investigated one method
of determining the relative importance of features: the correlation coefficient.
This leaves much room to use other ways of ranking features, such as the
information gain and 1R accuracy (which we have used in our work as feature
selectors).

Another area of potential future work is to investigate \textit{ensembles} of
classifiers, which take the predictions of more than one classifier (of the
same or different kind) and combine them to make a final prediction. There are
numerous ways in which this can be done: an example is to take the majority
class that is predicted as the final prediction for our ensemble. Ensembles
have often been demonstrated to perform better than any single classifier,
and there is a wide range of methods and combinations we could evaluate in
order to assess their performance relative to the single classifiers we have
used in our work.