### 2014-11-03 thesis/thanks/thanks.tex

\chapter*{Acknowledgements}

I would like to thank my supervisor, Dr Irena Koprinska, for her unwavering
support throughout the year. Irena, thank you for always being willing to
listen, and for raising points that inspired me to pursue new directions in
my research.

I would also like to thank Dr Michael Dinh from the Royal Prince Alfred
Hospital, without whom the research in this thesis would not have been
possible. Thank you for providing access to the data we needed for this study,
and for patiently answering all my questions about the data set and your
previous experiments.

Thank you also to Professor Paulo Cortez from the University of Minho in
Portugal for providing an extensive, independent data set from all the way
across the world.

Thank you to all of my friends, in particular Daniel Buckmaster, Stephen
Caraher and Christine Au Yeung, for the selfless encouragement, help and good
times which have made this year-long journey possible.

And, last but not least -- to my parents Caixiu and Ning, and
my sister Lily -- thank you for tirelessly supporting me throughout this year,
especially towards the deadline. This would not have been possible without all
of you.