### 2014-11-03 thesis/results/features-selected-nothreshold.tex
Add tables from feature selection

\begin{table}[htbp]
\resizebox{\textwidth}{!}{%
\tabcolsep=2pt
\begin{tabular}{llcccccccc}
\multicolumn{1}{c}{} & \multicolumn{1}{c}{} & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
\hline
 &  & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
 Type & Method & \multicolumn{1}{l}{Selected} & \multicolumn{1}{l}{\% of total} & \multicolumn{1}{l}{Selected} & \multicolumn{1}{l}{\% of total} & Selected & \% of total & Selected & \% of total \\
 \hline
 \multirow{2}{*}{Manual} & Baseline & 19 & 24.36 & - & - & 19 & 24.36 & - & - \\
  & Expert Selection & 11 & 14.10 & - & - & 11 & 14.10 & - & - \\
  \multirow{2}{*}{Automatic} & CFS & 9 & 11.54 & 2 & 14.29 & 9 & 11.54 & 2 & 14.29 \\
   & C4.5 Wrapper & 10 & 12.82 & 4 & 28.57 & 11 & 14.10 & 4 & 28.57 \\
   \end{tabular}
}
\caption{A listing of the number of features selected by non-threshold feature selectors for both data sets, with and without discretisation.}
\label{tab:features-selected-nothreshold}
\end{table}