### 2014-11-03 thesis/results/features-selected-threshold.tex

\begin{table}[htbp]
\resizebox{\textwidth}{!}{%
\begin{tabular}{lcccccccc}
\multicolumn{9}{c}{Correlation} \\
 & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
 \hline
  & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
  Threshold & Selected & \% of total & Selected & \% of total & Selected & \% of total & Selected & \% of total \\
  \hline
  0 & 58 & 74.36 & 14 & 100.00 & 57 & 73.08 & 13 & 92.86 \\
  0.01 & 55 & 70.51 & 14 & 100.00 & 54 & 69.23 & 13 & 92.86 \\
  0.05 & 41 & 52.56 & 12 & 85.71 & 41 & 52.56 & 12 & 85.71 \\
  0.1 & 27 & 34.62 & 10 & 71.43 & 29 & 37.18 & 10 & 71.43 \\
  0.125 & 19 & 24.36 & 7 & 50.00 & 20 & 25.64 & 7 & 50.00 \\
  0.15 & 15 & 19.23 & 6 & 42.86 & 12 & 15.38 & 6 & 42.86 \\
  0.2 & 9 & 11.54 & 4 & 28.57 & 7 & 8.97 & 4 & 28.57 \\
  0.3 & 2 & 2.56 & 1 & 7.14 & 1 & 1.28 & 1 & 7.14 \\
  0.5 & 0 & 0.00 & 1 & 7.14 & 0 & 0.00 & 1 & 7.14 \\
  0.75 & 0 & 0.00 & 1 & 7.14 & 0 & 0.00 & 1 & 7.14 \\
   &  &  &  &  &  &  &  &  \\
   \multicolumn{9}{c}{IG} \\
    & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
    \hline
     & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
     Threshold & Selected & \% of total & Selected & \% of total & Selected & \% of total & Selected & \% of total \\
     \hline
     0 & 57 & 73.08 & 13 & 92.86 & 57 & 73.08 & 14 & 100.00 \\
     0.001 & 47 & 60.26 & 13 & 92.86 & 47 & 60.26 & 14 & 100.00 \\
     0.005 & 33 & 42.31 & 12 & 85.71 & 33 & 42.31 & 12 & 85.71 \\
     0.01 & 27 & 34.62 & 12 & 85.71 & 27 & 34.62 & 12 & 85.71 \\
     0.05 & 5 & 6.41 & 8 & 57.14 & 4 & 5.13 & 8 & 57.14 \\
     0.1 & 2 & 2.56 & 7 & 50.00 & 2 & 2.56 & 7 & 50.00 \\
     0.15 & 1 & 1.28 & 5 & 35.71 & 1 & 1.28 & 5 & 35.71 \\
     0.2 & 0 & 0.00 & 3 & 21.43 & 0 & 0.00 & 3 & 21.43 \\
     0.3 & 0 & 0.00 & 2 & 14.29 & 0 & 0.00 & 2 & 14.29 \\
     0.5 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00 \\
      &  &  &  &  &  &  &  &  \\
      \multicolumn{9}{c}{1R} \\
       & \multicolumn{4}{c}{No discretisation} & \multicolumn{4}{c}{Discretisation} \\
       \hline
        & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} & \multicolumn{2}{c}{Trauma} & \multicolumn{2}{c}{General hospital} \\
        Threshold & Selected & \% of total & Selected & \% of total & Selected & \% of total & Selected & \% of total \\
        \hline
        66.905 & 55 & 70.51 & 14 & 100.00 & 59 & 75.64 & 14 & 100.00 \\
        67.905 & 4 & 5.13 & 14 & 100.00 & 4 & 5.13 & 14 & 100.00 \\
        68.905 & 3 & 3.85 & 14 & 100.00 & 2 & 2.56 & 14 & 100.00 \\
        69.905 & 3 & 3.85 & 14 & 100.00 & 2 & 2.56 & 14 & 100.00 \\
        71.905 & 1 & 1.28 & 14 & 100.00 & 0 & 0.00 & 14 & 100.00 \\
        74.405 & 0 & 0.00 & 14 & 100.00 & 0 & 0.00 & 14 & 100.00 \\
        76.905 & 0 & 0.00 & 6 & 42.86 & 0 & 0.00 & 6 & 42.86 \\
        81.905 & 0 & 0.00 & 3 & 21.43 & 0 & 0.00 & 3 & 21.43 \\
        86.905 & 0 & 0.00 & 2 & 14.29 & 0 & 0.00 & 2 & 14.29 \\
        96.905 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00 & 0 & 0.00
        \end{tabular}
}
\caption{A listing of the number of features selected by threshold for correlation, information gain and 1R for both data sets, with and without discretisation.}
\label{tab:features-selected-threshold}
\end{table}