### 2014-11-03 thesis/method/feature-descriptions.tex

\renewcommand{\arraystretch}{0.7}
\begin{longtable}{lp{13cm}}
\textbf{Name} & \textbf{Description} \\ 
\hline
id & The id of the current record (numeric) \\ 
age & Age of patient in years (numeric) \\ 
sex & Patient gender (nominal) \\ 
bp & Systolic blood pressure in mmHg (numeric) \\ 
bp1 & Whether or not the patient had a normal systolic blood pressure between 90-180 mmHg (nominal) \\ 
pr1 & Whether or not the patient had a normal pulse rate between 50-110 BPM (nominal) \\ 
rr1 & Whether or not the patient had a normal respiratory rate of 10-24 breaths per minute (nominal) \\ 
normalvitals & Whether or not the patient had bp, rr and pr normal (nominal) \\ 
pr & Pulse rate in BPM (numeric) \\ 
rr & Respiratory rate in breaths per minute (numeric) \\ 
gcs1 & Whether or not the patient had a Glasgow Coma Score of 14 or 15 (nominal) \\ 
gcs & Glasgow Coma Score on arrival (numeric) \\ 
iculos & Length of ICU stay in days (numeric) \\ 
icu & Whether or not the patient stayed in the ICU (nominal) \\ 
outcome & Whether the patient survived or died (nominal) \\ 
disp\_from\_ed & Disposition from hospital (nominal) \\ 
disp\_from\_hosp & Discharge destination from hospital (nominal) \\ 
iss8 & Whether or not the patient had an Injury Severity Score 8 or less (nominal) \\ 
iss & Injury Severity Score (numeric) \\ 
mechanism & Mechanism of injury (nominal) \\ 
male & Whether or not the patient was male (nominal) \\ 
age65 & Whether or not the patient was older than 65 years (nominal) \\ 
arrivalmode & How the patient arrived at the hospital (nominal) \\ 
transfer & Whether or not the patient was transferred from another hospital (nominal) \\ 
traumatier & The level of trauma team activation (nominal) \\ 
traumateam & Whether or not the trauma team was required (nominal) \\ 
penetrating & Whether or not the injury was penetrating (nominal) \\ 
died & Whether or not the patient died in hospital (nominal) \\ 
eddc & Patient was discharged within expected time (nominal) \\ 
rehab & Whether or not the patient was discharged from hospital to rehabiliation (nominal) \\ 
mechcode & Category of injury mechanism (nominal) \\ 
fall & Mechanism of injury was a fall (nominal) \\ 
headany & Whether or not there was any head injury (nominal) \\ 
faceany & Whether or not there was any face injury (nominal) \\ 
headface & Presence of any injury to head or face (nominal) \\ 
neckany & Any injury to neck (nominal) \\ 
chestany & Any injury to chest (nominal) \\ 
abdoany & Any injury to abdomen (nominal) \\ 
spineany & Any injury to spine (nominal) \\ 
upperlimbany & Any injury to upper limb (nominal) \\ 
lowerlimbany & Any injury to lower limb (nominal) \\ 
bodyregions & Number of bodyregions injured (numeric) \\ 
externalany & Any external burns injury (nominal) \\ 
head3 & Any major head injury (nominal) \\ 
face3 & Any major face injury (nominal) \\ 
neck3 & Any major neck injury (nominal) \\ 
chest3 & Any major chest injury (nominal) \\ 
abdo3 & Any major abdominal injury (nominal) \\ 
spine3 & Any major spine injury (nominal) \\ 
upper3 & Any major upper limb injury (nominal) \\ 
lower3 & Any major lower limb injury (nominal) \\ 
ext3 & Any major external injury (nominal) \\ 
laparotomy & Whether a laparotomy was performed (nominal) \\ 
thoracotomy & Whether a thoracotomy was performed (nominal) \\ 
neurosurgery & Whether any neurosurgery was performed (nominal) \\ 
anypelvis & Any injury to pelvic region (nominal) \\ 
pelvic\_fracture & Whether or not there was any pelvic fracture (nominal) \\ 
upper\_limbop & Whether an operation was performed on the upper limb (nominal) \\ 
lower\_limbop & Whether an operation was performed on the lower limb (nominal) \\ 
married & Whether or not the patient is married (nominal) \\ 
english & Whether or not the patient speaks English (nominal) \\ 
charlson & Charlson Comorbidity Index (numeric) \\ 
triss & Calculated probability of survival (numeric) \\ 
mentalhealth & Whether the patient suffered any mental health problems (nominal) \\ 
alcohol & Whether alcohol was involved in the injury (nominal) \\ 
comorbidity & Whether or not the patient had comorbidities (nominal) \\ 
operation & Whether or not an operation was performed in the hospital (nominal) \\ 
upperlimboponly & An operation was performed only on the upper limb (nominal) \\ 
upperlimbonly & Any injury to the upper limb only (nominal) \\ 
headminor & Minor or no head injury (nominal) \\ 
lowerlimbnopelvis & Any injury to the lower limb, excluding pelvic injuries (nominal) \\ 
any\_heart & Any heart disease (nominal) \\ 
any\_liver & Any liver disease (nominal) \\ 
any\_kidney & Any kidney disease (nominal) \\ 
any\_cancer & Whether the patient has or had cancer (nominal) \\ 
any\_dm & Any diabetes mellitus (nominal) \\ 
any\_stroketia & Whether or not the patient had suffered a stroke or transient ischemic attack (nominal) \\ 
los & Length of hospital stay in days (numeric) \\ 
los48 & Whether or not the length of hospital stay was less than or equal to 2 days (nominal) \\ 
\caption{A listing of all features from the trauma data set and what they represent.}
\label{tab:feature-desc-tr}
\end{longtable}
\renewcommand{\arraystretch}{1}

\begin{table}[htbp]
\begin{tabular}{ll}
\textbf{Name} & \textbf{Description} \\ 
\hline
Sex & Patient gender (nominal) \\ 
Age & Age at the time of admission, separated into categories (nominal) \\ 
Education & Highest level of education completed (nominal) \\ 
Marital Status & Marital status (nominal) \\ 
Episode Type & Patient type of episode (nominal) \\ 
Inpatient Service & Physical inpatient service (nominal) \\ 
Medical Specialty & Medical specialty associated with patient condition (nominal) \\ 
Origin Episode Type & Origin episode type that caused hospitalisation (nominal) \\ 
Admission Month & Month admitted to hospital (nominal) \\ 
Admission Day & Day of week admitted to hospital (nominal) \\ 
Admission Hour & Hour of the day patient was admitted to hospital (nominal) \\ 
Main Procedure & Type of main procedure performed (nominal) \\ 
Main Diagnosis & Main diagnosis of patient (nominal) \\ 
Previous Admissions & Number of times patient was previously admitted (numeric) \\ 
LOS48 & Whether or not the length of hospital stay was less than or equal to 2 days (nominal) \\ 
\end{tabular}
\caption{A listing of all features from the general hospital data set and what they represent.}
\label{tab:feature-desc-pt}
\end{table}