### 2014-11-03 thesis/results/overall-portugal-auc.tex

\begin{table}[htbp]
\begin{tabular}{l|cccccc}
Classifier & None & CFS & C4.5 Wrapper & Correlation & IG & 1R \\ \hline
ZeroR & 0.5* & 0.5* & 0.5* & 0.5* & 0.5* & 0.5* \\
1R & 0.858* & 0.858* & 0.858* & 0.858* & 0.858* & 0.858* \\
NB & \begin{tabular}[c]{@{}c@{}}0.979\\ D\end{tabular} & 0.982* & \textbf{0.983*} & \begin{tabular}[c]{@{}c@{}}0.981*\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.979*\\ t=74.405\end{tabular} \\
DT & \textbf{0.989} & 0.981* & 0.984* & \begin{tabular}[c]{@{}c@{}}0.989\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989*\\ t=0.001\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989*\\ t=74.405\end{tabular} \\
LR & 0.994* & 0.981* & 0.990* & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=74.405\end{tabular} \\
SVM & 0.994* & 0.971* & 0.981* & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.01\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=74.405\end{tabular} \\
1NN & \begin{tabular}[c]{@{}c@{}}0.979\\ D\end{tabular} & 0.982* & \textbf{0.991*} & \begin{tabular}[c]{@{}c@{}}0.991*\\ t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.989*\\ t=76.905\end{tabular} \\
20NN & 0.992* & 0.981* & 0.991* & \textbf{\begin{tabular}[c]{@{}c@{}}0.992*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.992*\\ t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.992*\\ t=74.405\end{tabular} \\
RD & 0.993* & 0.794* & 0.986* & \textbf{\begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.005\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.993*\\ t=74.405\end{tabular} \\
K* & 0.994 & 0.980* & 0.989* & \begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}0.994*\\ t=0.05\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.994\\ t=74.405\end{tabular} \\
MLP & 0.990* & 0.983* & 0.992* & \textbf{\begin{tabular}[c]{@{}c@{}}0.993*\\ t=0.15\end{tabular}} & \begin{tabular}[c]{@{}c@{}}0.991*\\ t=0.1\end{tabular} & \begin{tabular}[c]{@{}c@{}}0.990*\\ t=76.905\end{tabular} \\
\end{tabular}
\caption{Overall best AUC of each classifier with a particular feature selection method on the general hospital data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\label{tab:overall-pt-auc}
\end{table}