### 2014-11-04 thesis/appendix/code.tex

\chapter{Source Code Listings}
\label{app:code}

In this appendix we display the source code of the most important programs we
wrote in our to conduct our experiments. These are described in more detail
in Section \ref{sec:code}. To download the latest version of the code, please
visit \url{https://github.com/tianyupu/hons-thesis}.

\lstset{
  language = Java,
  texcl = false, % Parse comments as TeX
  backgroundcolor = \color{white},
  basicstyle = \footnotesize\ttfamily\singlespacing,
  commentstyle = \color{mygreen},
  keywordstyle = \color{blue},
  stringstyle = \color{mymauve},
  numberstyle = \tiny\color{mygray},
  numbers = left, % possible values are (none, left, right)
  numbersep = 5pt, % how far the line-numbers are from the code
  rulecolor = \color{black},
  stepnumber = 1, % the step between two line-numbers
  showstringspaces = false, % underline spaces within strings?
  literate = {~} {$\sim$}{1}, % make ~ nice
  literate = {_} {\textunderscore}{1}, % make _ nice
  breaklines = true,
  prebreak = \raisebox{0ex}[0ex][0ex]
  {\color{red}\ensuremath{\space\hookleftarrow}},
  postbreak = \raisebox{0ex}[0ex][0ex]
  {\color{red}\ensuremath{\hookrightarrow\space}},
  tabsize = 2
}

\section{Ranked Distance Implementation}
\label{sec:rankeddist-code}
\lstset{language=Java}
\lstinputlisting{../code/weka/RankedDistance.java}

\section{Data Cleaning and Whitespace Removal}
\lstset{language=Python}
\label{sec:python-preprocess}
\lstinputlisting{../code/sklearn/util.py}

\section{Automating WEKA and Saving Results}
\label{sec:run-weka}
\lstset{language=bash}
\lstinputlisting{../code/weka/run_weka.sh}

\subsection{WEKA Configuration File}
The configuration files that were used in the above script consisted of WEKA
classes in a text file, one per line with as many as we wanted to test in one
experiment. A typical file looked like this:
\lstset{language={}}
\lstinputlisting{../code/weka/weka-config/all_clfs_default}

\section{Summarising a Large Number of Results}
\label{sec:summarise-script}
\lstset{language=bash}
\lstinputlisting{../code/weka/summarise_results.sh}

\subsection{Computing Confidence Intervals from Summarised Results}
\label{sec:compute-ci}
\lstset{language=Python}
\lstinputlisting{../code/weka/compute_ci.py}