### 2014-11-07 presentation/slides.tex
Add skeleton slides

\documentclass[compress]{beamer}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{verbatim}

% Load the usyd template
\input{usyd-beamer/usyd-beamer.tex}

\begin{document}
\usydtitleframe{Data Mining for Predicting Length of Hospital Stay}{}{Tianyu Pu \\
  Supervisor: Irena Koprinska}

\section{Introduction}
\subsection{}
\begin{frame}{Thought exercise}
Imagine that you have just been admitted to hospital...
\end{frame}

\subsection{}
\begin{frame}{Outline of talk}
The problem

Previous work

Our approach

Results

Conclusions and future work
\end{frame}

\section{The Problem}
\subsection{}
\begin{frame}{The problem}
\end{frame}

\section{Background}
\subsection{}
\begin{frame}{Common approach: logistic regression}
\end{frame}

\subsection{}
\begin{frame}{Artificial neural networks}
\end{frame}

\subsection{}
\begin{frame}{Feature selection}
\end{frame}

\section{Approach}
\subsection{}
\begin{frame}{Ranked Distance Nearest Neighbour}
\end{frame}

\subsection{}
\begin{frame}{Overview of our approach}
\end{frame}

\section{Results}
\subsection{}
\begin{frame}{Trauma data set}
\end{frame}

\subsection{}
\begin{frame}{General hospital data set}
\end{frame}

\section{Conclusions}
\subsection{}
\begin{frame}{Conclusions}
\end{frame}

\end{document}