### 2014-12-23 presentation/summer-slides.tex
Add slides for summer scholarship presentation

\documentclass[compress]{beamer}
\usepackage{epstopdf}
\usepackage{graphicx}
\usepackage{lmodern}
\usepackage{tikz}
\usepackage{verbatim}

\usetikzlibrary{decorations.pathreplacing,calc}

\newcommand{\tikzmark}[2][-3pt]{\tikz[remember picture, overlay, baseline=-2.0ex]\node[#1](#2){};}

\tikzset{brace/.style={decorate, decoration={brace}},
 brace mirrored/.style={decorate, decoration={brace,mirror}},
}

\newcounter{brace}
\setcounter{brace}{0}
\newcommand{\drawbrace}[3][brace]{%
 \refstepcounter{brace}
  \tikz[remember picture, overlay]\draw[#1] (#2.center)--(#3.center)node[pos=0.5, name=brace-\thebrace]{};
}

\newcounter{arrow}
\setcounter{arrow}{0}
\newcommand{\drawcurvedarrow}[3][]{%
 \refstepcounter{arrow}
  \tikz[remember picture, overlay]\draw (#2.center)edge[#1]node[coordinate,pos=0.5, name=arrow-\thearrow]{}(#3.center);
}

% #1 options, #2 position, #3 text 
\newcommand{\annote}[3][]{%
 \tikz[remember picture, overlay]\node[#1] at (#2) {#3};
}

\renewcommand*\footnoterule{} % remove footnote rule

\definecolor{emerald}{RGB}{0,166,79}

% Load the usyd template
\input{usyd-beamer/usyd-beamer.tex}

\setbeamerfont{footnote}{size=\tiny}

\begin{document}
\usydtitleframe{Predicting the Length of Stay in Hospital Using Data Mining Methods}{}{\textbf{Tianyu Pu} \\
  Supervisors: Irena Koprinska \& Kalina Yacef}

\section{Introduction}
\subsection{}
\begin{frame}{Introduction}
Imagine that you have just been admitted to hospital...
  \pause

\vspace{0.5cm}
\textbf{Why do we care?}
\end{frame}

\section{The Problem}
\subsection{}
\begin{frame}{The problem}
Given
\begin{center}
\begin{tabular}{cccc|c}
\tikzmark[]{x}age & gender & heart rate & ... & length of stay (LOS)\tikzmark[]{y} \\
\hline
\tikzmark[xshift=-0.80ex, yshift=-1ex]{a}45 & F & 122 & ... & $>$ 2 days \\
23 & M & 98 & ... & $\leq$ 2 days \\
\multicolumn{4}{c|}{\tikzmark[xshift=-13.3ex, yshift=-2ex]{b}...} & ... \\
\end{tabular}
\end{center}
\drawbrace[brace mirrored, thick, gray]{y}{x}
\drawbrace[brace mirrored, thick, gray]{a}{b}
\annote[yshift=1.5ex, gray]{brace-1}{\scriptsize features}
\annote[left, gray]{brace-2}{\scriptsize examples}

\vspace{0.5cm}
Predict LOS category for new patient
\begin{center}
\begin{tabular}{cccc|c}
age & gender & heart rate & ... & length of stay (LOS)  \\
\hline
34 & F & 101 & ... & ? \\
\end{tabular}
\end{center}
\end{frame}

\section{Background and Previous Work}
\subsection{}
\begin{frame}{Prediction}
\begin{center}
\includegraphics<1>{learning.pdf}
\includegraphics<2>{learning-focus2.pdf}
\end{center}
\end{frame}

\subsection{}
\begin{frame}{Feature discretisation and selection}
\textbf{Discretisation}
\begin{itemize}
\item Convert numeric features into categorical features
\item \textit{Supervised} discretisation
\end{itemize}

\vspace{0.5cm}

  \pause
\textbf{Selection}
\begin{itemize}
\item Find a subset of features that are the most `relevant'
\item Two ways: \textit{manual} and \textit{automatic}
\end{itemize}
\end{frame}

\subsection{}
\begin{frame}{Learning algorithms}
\textbf{Logistic regression}

One of the most widely used approaches in LOS prediction
\begin{equation*}
\begin{aligned}
\text{log}\left(\dfrac{\text{probability of LOS}\leq\text{2 days}}{\text{probability of LOS}>\text{2 days}}\right) &= b_0 + b_1\text{age} + b_2\text{gender} + \ldots
\end{aligned}
\end{equation*}
  \pause

\textbf{Others}

Artificial neural networks, ...
\end{frame}

\subsection{}
\begin{frame}{Where We're At}
Comprehensive evaluation of feature selection, discretisation and learning algorithms
\pause

\vspace{0.5cm}
Proposed new nearest-neighbour approach
\end{frame}

\section{Next Steps}
\subsection{}
\begin{frame}{Next Steps}
Feedback from domain experts

\pause
\vspace{0.5cm}
Draft papers for publication
\pause

\vspace{0.5cm}
Investigate ensembles and refinement of nearest-neighbour method
\end{frame}

\end{document}