### 2015-03-07 papers/results/overall-trauma-acc.tex
Add overall tables and move plots into a separate file

\begin{table}[htbp]
\resizebox{\linewidth}{!}{%
\begin{tabular}{r|cccccc}
Classifier & None & Baseline & Expert & C4.5 Wrapper & Correlation & IG \\ \hline
LR & \begin{tabular}[c]{@{}c@{}}77.62 (0.09)\\ D\end{tabular} & 75.06 (0.06)* & \begin{tabular}[c]{@{}c@{}}76.45 (0.06)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.3 (0.06)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.7 (0.09)\\ D, t=0\end{tabular}} & \begin{tabular}[c]{@{}c@{}}77.65 (0.08)\\ D, t=0\end{tabular} \\
SVM & \begin{tabular}[c]{@{}c@{}}77.39 (0.08)\\ D\end{tabular} & 74.47 (0.07)* & \begin{tabular}[c]{@{}c@{}}75.67 (0.02)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}75.26 (0.07)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}77.42 (0.06)\\ D, t=0.1\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.53 (0.07)\\ D, t=0.005\end{tabular}} \\
1NN & \begin{tabular}[c]{@{}c@{}}71.22 (0.1)\\ D\end{tabular} & 70.28 (0.08)* & \begin{tabular}[c]{@{}c@{}}72.56 (0.11)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.81 (0.07)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.75 (0.13)\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}73.26 (0.13)\\ D, t=0.01\end{tabular} \\
20NN & \begin{tabular}[c]{@{}c@{}}75.9 (0.13)\\ D\end{tabular} & 73.59 (0.15)* & \begin{tabular}[c]{@{}c@{}}75.51 (0.19)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.48 (0.13)\\ D\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.26 (0.2)\\ D, t=0.125\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.56 (0.18)\\ D, t=0.005\end{tabular}} \\
RD & \begin{tabular}[c]{@{}c@{}}76.92 (0.11)\\ D\end{tabular} & 69.68 (0.3)* & \begin{tabular}[c]{@{}c@{}}76.19 (0.16)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.22 (0.14)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}76.87 (0.25)\\ D, t=0.05\end{tabular} & \begin{tabular}[c]{@{}c@{}}76.82 (0.16)\\ D, t=0.001\end{tabular} \\
K* & \begin{tabular}[c]{@{}c@{}}71.4 (0.1)\\ D\end{tabular} & 72.20 (0.11)* & \begin{tabular}[c]{@{}c@{}}75.08 (0.09)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}77.27 (0.07)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}75.07 (0.14)\\ D, t=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.08 (0.12)\\ D, t=0.01\end{tabular} \\
MLP & \begin{tabular}[c]{@{}c@{}}73.79 (0.16)\\ D\end{tabular} & 73.03 (0.15)* & \begin{tabular}[c]{@{}c@{}}73.45 (0.16)\\ D\end{tabular} & \textbf{\begin{tabular}[c]{@{}c@{}}76.66 (0.12)\\ D\end{tabular}} & \begin{tabular}[c]{@{}c@{}}74.65 (0.17)\\ T=0.15\end{tabular} & \begin{tabular}[c]{@{}c@{}}74.03 (0.18)\\ D, t=0.001\end{tabular}
\end{tabular}
}
\caption{Overall best accuracy of each classifier with a particular feature selection method on the trauma data set. * indicates that results were identical regardless of discretisation. D indicates that discretisation was used, and $t$ is the threshold at which the feature selector was applied. The absence of * or D means that not discretising produced a better result. The best result for each classifier is highlighted in bold, with ties broken in favour of the smallest feature set.}
\label{tab:overall-tr-acc}
\end{table}