### 2015-03-07 papers/results/tr-corr-ig-plots.tex

\begin{figure}[htbp]
\begin{subfigure}{1.1\textwidth}
\includegraphics[width=\textwidth]{results/tr-corr.eps}
\caption{}
\label{fig:tr-threshold-corr}
\end{subfigure}

\begin{subfigure}{1.1\textwidth}
\includegraphics[width=\textwidth]{results/tr-ig.eps}
\caption{}
\label{fig:tr-threshold-ig}
\end{subfigure}
\caption[]{Comparison of accuracy and AUC between classifiers for feature selectors that discard features based on a threshold. Solid lines correspond to a result without discretisation, and dashed lines correspond to a result with discretisation.}
\label{fig:tr-threshold}
\end{figure}