### 2015-03-07 papers/results/tr-nothreshold-plots.tex

\begin{figure*}[htbp]
\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-nofs-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-nofs-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-dinh-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-dinh-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-expert-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-expert-auc.eps}
\caption{}
\label{}
\end{subfigure}

\begin{subfigure}{.50\textwidth}
\includegraphics[width=\textwidth]{results/tr-wrapper-acc.eps}
\caption{}
\label{}
\end{subfigure}%
\begin{subfigure}{.57\textwidth}
\includegraphics[width=\textwidth]{results/tr-wrapper-auc.eps}
\caption{}
\label{}
\end{subfigure}
\caption[]{Comparison of accuracy and AUC between classifiers for each non-threshold feature selection method, grouped by whether or not discretisation was applied.}
\label{}
\end{figure*}